### 2024-01-29 r/README.qmd
Update readme

---
format: gfm
---

```{r}
devtools::install()
```

For local development run this in the `r` directory:

```{r}
devtools::load_all()
```

```{r}
library(dplyr)
```

Basic test

```{r}
x = sf::read_sf("data-raw/geojson/princes_street_minimal.geojson")
y = sf::read_sf("data-raw/geojson/princes_street_minimal_x_1.geojson")
```

```{r}
bind_rows(
  princes_street_minimal |>
    transmute(value = as.character(value)),
  princes_street_minimal_x_1 |>
    transmute(value = "x")
) |>
  plot(lwd = 5)
```