### 2024-01-30 r/data-raw/test-large-edinburgh.qmd
Add large example, close #13

---
format: gfm
---

```{r}
open_roads = sf::read_sf("https://github.com/nptscot/networkmerge/releases/download/v0.1/os_3km.geojson")
summary(duplicated(open_roads$geometry))
plot(open_roads$geometry)
open_roads = sf::st_make_valid(open_roads)
```

We can get some OSM data of the same area as follows:

```{r}
case_study_area = open_roads |>
  sf::st_union() |>
  sf::st_convex_hull() |>
  sf::st_transform("EPSG:4326")
osm_roads = osmextract::oe_get_network(
  "Edinburgh",
  mode = "driving",
  boundary = case_study_area,
  boundary_type = "clip"
)
plot(osm_roads$geometry)
```

```{r}
table(osm_roads$highway)
```

```{r}
osm_roads$value = 1
osm_roads$value[osm_roads$highway == "primary"] = 10
osm_roads$value[osm_roads$highway == "secondary"] = 9
osm_roads$value[osm_roads$highway == "tertiary"] = 8
osm_roads$value[osm_roads$highway == "unclassified"] = 7
osm_roads$value[osm_roads$highway == "residential"] = 2
```

```{r}
funs = list(value = sum)
```

```{r}
os_joined = stplanr::rnet_merge(
  open_roads |> dplyr::transmute(n = seq(nrow(open_roads))) |> sf::st_transform("EPSG:27700"),
  osm_roads |> sf::st_cast("LINESTRING") |> sf::st_zm() |> sf::st_transform("EPSG:27700"),
  funs = funs,
  dist = 20,
  segment_length = 10
  )
os_joined |>
  dplyr::select(value) |>
  plot(logz = TRUE)
```